Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-18 E-Hypergraphs-LICS24.tex
proofs for duality; pattern matching references

\begin{abstract}
The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
We give a categorical semantics to these structures, called \emph{e-graphs}, in terms of Cartesian categories enriched over a semilattice. 
We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
Finally, we present a sound and complete combinatorial representation of morphisms in such a category, based on a generalisation of hypergraphs (\emph{e-hypergraphs}).
They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
\end{abstract}

>>>>

\begin{abstract}
The technique of equipping graphs with an equivalence relation, called equality saturation, has recently proved both powerful and practical in program optimisation, particularly for SMT solvers. 
We give a categorical semantics to these structures, called e-graphs, in terms of Cartesian categories enriched over a semilattice. 
We show how this semantics can be generalised to monoidal categories, which opens the door to new applications of e-graph techniques, from algebraic to monoidal theories.
Finally, we present a sound and complete combinatorial representation of morphisms in such a category,  based on a generalisation of hypergraphs which we call e-hypergraphs.
They have the usual advantage that many of their structural equations are absorbed into a general notion of isomorphism. 
\end{abstract}
